---
title: "SPL validation on the PBC cohort"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SPL validation on the PBC cohort}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette illustrates SPL scoring on the primary biliary cirrhosis cohort available in the `survival` package. The code below reproduces the model-by-kernel comparison table used in the real-data application.

```{r setup, message=FALSE}
library(splcr)
library(survival)
```

## Data preparation

The package assumes that follow-up time is recorded in **years**.


```{r}
data(pbc, package = "survival")

pbc_data <- subset(pbc, !is.na(trt))
pbc_data$status <- ifelse(pbc_data$status == 2, 1L, ifelse(pbc_data$status == 1, 2L, 0L))
pbc_data$time <- as.numeric(pbc_data$time) / 365.25

# Transformations consistent with the application code
pbc_data$logbili <- log(pbc_data$bili)
pbc_data$sex <- factor(pbc_data$sex)
pbc_data$stage <- factor(pbc_data$stage)
pbc_data$ascites <- factor(pbc_data$ascites)

keep <- c("time","status","age","sex","stage","logbili","albumin","protime","ascites")
pbc_data <- pbc_data[complete.cases(pbc_data[, keep]), keep]
```

## Candidate models

```{r}
models <- list(
  "M1: Baseline Clinical"    = ~ age + sex + stage,
  "M2: Core Biochemical"     = ~ logbili + albumin,
  "M3: Extended Biochemical" = ~ logbili + albumin + protime + ascites,
  "M4: Clinical + Core Bio"  = ~ age + sex + stage + logbili + albumin,
  "M5: Full"                 = ~ age + sex + stage + logbili + albumin + protime + ascites
)

kernels <- c("uniform","triangular","epanechnikov","quartic")
seed <- 20250101
```

## SPL-CSH table

```{r, eval=FALSE}
tab_csh <- spl_compare(models, pbc_data, method = "csh",
                      folds = 10, k_nn = 5,
                      kernels = kernels,
                      seed = seed)

tab_csh
```

## SPL-FG table

```{r, eval=FALSE}
tab_fg <- spl_compare(models, pbc_data, method = "fg",
                      folds = 10, k_nn = 5,
                      kernels = kernels,
                      seed = seed)

tab_fg
```

The two chunks above are marked as `eval = FALSE` because they fit multiple models across several folds and kernels. To run them locally, set `eval = TRUE` and re-knit the vignette.
